---
title: "Feb27Class"
author: "Sokona Mangane"
date: '2023-02-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
#install.packages("simEd")
install.packages("magick")
library(simEd)
library(magick)
```



ctl + shift + c = comment
ctrl + Alt + i = r code chunk
ctrl + shift + m = pipe operator

# Evemt - Driven Simulation

Recall the time-persistent statistics:
* $n(t)$, the number of jobs in the system
* $q(t)$, the number of jobs in the queue
* $x(t)$, the number of jobs in service

▶ State: $n(t)$ provides a complete characterization
$n(t) = 0$ ⇐⇒ $q(t) = 0$ and $x(t) = 0$
$n(t) > 0$ ⇐⇒ $q(t) = n(t) − 1$and $x(t) = 1$


▶ Event types:
* arrival: increases $n(t)$ by 1
* completion: decreases $n(t)$ by 1


```{r}
ssqvis(maxArrivals = 50, seed = 8675309)

```

# In Class Work


```{}

initTimes <- function() {
     arrivalTimes      <<- c(15, 47, 71, 111, 123, 152, 232, 245, 99999)
     interarrivalTimes <<- c(arrivalTimes[1], diff(arrivalTimes))
     serviceTimes      <<- c(43, 36, 34, 30, 38, 30, 31, 29)
 }

 getInterarr <- function() {
     nextInterarr <- interarrivalTimes[1]
     interarrivalTimes <<- interarrivalTimes[-1] # remove 1st element globally
     return(nextInterarr)
 }

 getService <- function() {
     nextService <- serviceTimes[1]
     serviceTimes <<- serviceTimes[-1]  # remove 1st element globally
     return(nextService)
 }

initTimes()
output <- ssq(maxArrivals = 8, interarrivalFcn = getInterarr, serviceFcn = getService, saveAllStats = TRUE)




```
