---
title: "Feb6 Class"
author: "Sokona Mangane"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# rng

g(x) = ax mod m -> x/m will be uniformly distrubitued between (0,1)

# Random Variables

X is a random varialbe, written simply as X. Reminder: probabilites have to be greater than 0 but bigger than 1

PMF: Probability Mass Function: probability that X takes on the value x, *f(x) = Pr(X  = x)*.


```{r pressure, echo=FALSE}
par(mfrow = c(2,1))
xvals <- 0:10
plot(xvals, dgeom(xvals, prob = 0.5), type = "h") #geometric(0.5) pmf
plot(xvals, pgeom(xvals, prob = 0.5), type = "S") #geometric(0.5) cdf

```

random variate: aglorithmixally generated relaization of a random variable

```{r}
igeom(0.7, 0.1) # the 2nd graph is a PMF and the 3rd was the CDF, inverse distribution happening on the fly (geometrc distribution)
```


stationery poisson process = rate doesnt change, exponential interarrivals 

```{r}
numJobs <- 1000
maxT <- 50
  
times <- runif(numJobs, )
times = sort(times)
stripchart(times, pch = '|')
iatimes = diff(times)
hist(iatimes)
hist(iatimes, freq = FALSE, )
```





