---
title: "HW4"
author: "Sokona Mangane"
date: "2023-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Unifrom Random Variates:

1. Write an R function names `myunif` that accepts three parameters: n, the number of variates to generate; a, the lower limit of the distribution; and b, the upper limit of the distribution. 

Your function must return a vector of *n* uniform *(a, b)* random variates. Each variate must be generated by inverting a single call to runif(1) — use the idf determined by setting u = F (x) and solving for x (you are not permitted to use quniif).

```{r q1}

myunif <- function(n, a, b){
  values = rep(0,n)
  for(i in 0:n){
    values[i] = a + ((b-a)*(runif(1)))
  }
  return(values)
}

a = 3
b = 7
n = 100

hist(myunif(n, a, b))
```

2. Using your myunif function, produce a histogram of many uniform (a, b) random variates, superimposing the true theoretical distribution (see dunif).

```{r}
histdata = rep(0,n)
  for(i in 0:n){
    histdata[i] = a + ((b-a)*(runif(1)))
  }
  return(histdata)
histdata = 

```


3. Confirm that the mean and standard deviation of your random variate sample are consistent with the theoretical values given above

```{r q3}
tmean = (a + b)/2 #theoretical mean
tsd = b - a / sqrt(12) #theoretical sd

mean(histdata) == tmean
sd(histdata) == tsd


```



# Triangular Random Variates

Often, the only data known about a real-world distribution is its minimum value, its maximum value, and its most likely value (mode). In such a case, the triangular distribution is appropriate. Let the random variable X be triangular (a, b, c) where a is the minimum value, b is the maximum value, and c is the mode.

1. Derive the pdf f (x) for a triangular (a, b, c) random variable, where a < x < b. Hint: the function can be considered in two separate pieces: linear in (a, c), and separately linear in (c, b). **Show your work.**

```{r}

```

2. Using the pdf f (x) derived above, derive the cdf F (x). Again, the function will be separately linear in (a, c) and in (c, b). **Show your work.**

```{r}

```

3. Using the cdf F(x) derived above, derive the idf F−1(u) for 0 < u < 1. As before, derive by setting u = F(x) and solve for x. **Show your work.**

```{r}

```

4. Write an R function named mytriang that accepts four parameters: n, the number of variates to generate; a, the lower limit of the distribution; b, the upper limit of the distribution; and c, the mode of the distribution.

```{r}

```

Your function must return a vector of n triangular (a, b, c) random variates. Each variate must be generated by inverting a single call to runif(1) using the idf you determined above.

```{r}

```

5. Using your mytriang function, produce a histogram of many triangular (a, b, c) random variates, superimposing the true theoretical distribution (use the segments function in R to plot your derived f(x) equation from above).

```{r}

```

6. Confirm that the mean and standard deviation of your random variate sample are consistent with the theoretical values given below.


```{r pressure, echo=FALSE}
plot(pressure)
```

