---
title: "HW4"
author: "Sokona Mangane"
date: "2023-02-08"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Unifrom Random Variates:

1.  Write an R function names `myunif` that accepts three parameters: `n`, the number of variates to generate; `a`, the lower limit of the distribution; and `b`, the upper limit of the distribution.

Your function must return a vector of *n uniform (a, b)* random variates. Each variate must be generated by inverting a single call to `runif(1)` --- use the idf determined by setting $u = F (x)$ and solving for $x$ (you are not permitted to use `qunif`).

```{r q1}

myunif <- function(n, a, b){
  values = rep(0,n)
  for(i in 0:n){
    values[i] = a + ((b-a)*(runif(1)))
  }
  return(values)
}

a = 0
b = 1000
n = 1000

hist(myunif(n, a, b))
```

2.  Using your `myunif` function, produce a histogram of many *uniform (a, b)* random variates, superimposing the true theoretical distribution (see `dunif`).

```{r q2}

histdata = myunif(10000, 10, 20)
x <- seq(10,20, by = 0.01)
hist(histdata, freq = FALSE)
curve(dunif(x, 10, 20), add = TRUE)

```

3.  Confirm that the mean and standard deviation of your random variate sample are consistent with the theoretical values given above

```{r q3}
a = 10
b = 20
tmean = (a + b)/2 #theoretical mean
tsd = (b - a) / sqrt(12) #theoretical sd

mean(histdata) == tmean
sd(histdata) == tsd

```

Even though the values say FALSE, the mean is of our sample, `r mean(histdata)`, is a few points off, but very close to the theoretical mean, `r tmean`. The standard deviation of our sample, `r sd(histdata)`, is almost identical to the theoretical standard deviation of `r tsd`.

# Triangular Random Variates

Often, the only data known about a real-world distribution is its minimum value, its maximum value, and its most likely value (mode). In such a case, the *triangular* distribution is appropriate. Let the random variable X be *triangular (a, b, c)* where *a* is the minimum value, *b* is the maximum value, and *c* is the mode.

4.  Write an R function named `mytriang` that accepts four parameters: `n`, the number of variates to generate; `a`, the lower limit of the distribution; `b`, the upper limit of the distribution; and `c`, the mode of the distribution.

Your function must return a vector of *n triangular (a, b, c)* random variates. Each variate must be generated by inverting a single call to `runif(1)` using the idf you determined above.

```{r q4}

mytriang <- function(n, a, b, c){
  values = rep(0,n)
  f_c <- (c-a)/(b-a)
  for(i in 0:n){
    u = runif(1)
    if(0 < u && u < f_c) {
      d = (b - a)*(c - a)
      values[i] = a + (sqrt(d * u))
    } else if(f_c <= u && u < 1) {
      d = (b - a)*(b - c)*(1 - (u))
      values[i] = b - (sqrt(d))
    }
  }
  return(values)
}

a = 0
b = 1000
c = 500
n = 1000

hist(mytriang(n, a, b, c), freq = FALSE) 

```

5.  Using your `mytriang` function, produce a histogram of many *triangular (a, b, c)* random variates, superimposing the true theoretical distribution (use the segments function in R to plot your derived $f(x)$ equation from above).

```{r q5}

a = 10
b = 25
c = 15
n = 10000

histdata1 = mytriang(n, a, b, c)
hist(mytriang(n, a, b, c), freq = FALSE) 
segments(c, (2/(b-a)), b, 0)

```

6.  Confirm that the mean and standard deviation of your random variate sample are consistent with the theoretical values given below.

```{r q6}
ttriangmean = (a + b + c)/3 #theoretical triangular mean
ttriangsd = sqrt((1/18) * ((a^2) + (b^2) + (c^2) - (a*b) - (a*c) - (b*c))) #theoretical triangular sd

mean(histdata1) == ttriangmean
sd(histdata1) == ttriangsd


```

Even though the values say FALSE, the mean and the standard deviation of our sample, `r mean(histdata1)` and `r sd(histdata1)` respectively, are almost identical to the theoretical mean of `r ttriangmean` and theoretical standard deviation of `r ttriangsd` respectively. This means that our code and `mytriang` function are very close to the true theoretical triangular distribution.

