---
title: "Feb13Class"
author: "Sokona Mangane"
date: '2023-02-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown Class Notes

Ctrl + Alt + I = include code chunk
Ctrl + shift + m = include pipe operator

Galileo() can run two dice..?!

Monter Carlo Simulation Results:


* *More* replications doesn't guarantee a better estimate
* Frequency probability estimates converge slowly and somewhat erratically
* the period where ur graph is/points are moving around and erraticley is called the warmup period, which can't rlly be automated (you can't exactly where it starts or end)


```{r cars}
#Algorithm:
# wins = 0
# for ( i in 1:nrep ) {
#   roll = sample( 1:6, 1 ) + sample( 1:6, 1 )
#   if ( roll == 7 || roll == 11 ) {
#     wins = wins + 1
#     } else if ( roll != 2 && roll != 3 && roll != 12 ) {
#       point = roll
#       while ( TRUE ) {
#         roll = sample( 1:6, 1 ) + sample( 1:6, 1 )
#         if ( roll == point ) wins = wins + 1
#         if ( roll == point || roll == 7 ) break
#       }
#     }
# }
# 
# print( wins / nrep )
```

To get confidence intervals in R, you can use a function called t.test

95% confidencce interval means that we're 95% confident that the tru unknown mean lies within those bounds. So if you run experiments 200 times, than ~190  of those experiments will be within the tru theoritcal means. 

## In Class Work: Monte Carlo

*1. Implement your own function for craps, but simulating unfair dice, where each die is loaded in the following way:*
* *Each of 2, 3, 4, 5 are twice as likely as a 1.*
* *A 6 is four times as likely as a 1.*

```{r}

craps<- function(nrep = 100){
  list <- c(1, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, 6, 6)
  wins = 0
  for ( i in 1:nrep ) {
    roll = sample( list, 1 ) + sample( list, 1 ) #function that gives us a random die that 's loaded
    if ( roll == 7 || roll == 11 ) {
      wins = wins + 1
      } else if ( roll != 2 && roll != 3 && roll != 12 ) {
        point = roll
        while ( TRUE ) {
          roll = sample( list, 1 ) + sample( list, 1 )
          if ( roll == point ) wins = wins + 1
          if ( roll == point || roll == 7 ) break
        }
      }
  }
  
  return( wins / nrep )
}

```

*2. Execute your function 100, 1000, and 10 000 times, each time storing the results into a vector.*

```{r}

estimates100 <- rep(0,100)

estimates100 <- for (i in 1:100) {
  estimates100[i] <- craps()
}

estimates1000 <- rep(0,1000)

estimates1000 <- for (i in 1:1000) {
  estimates1000[i] <- craps()
}

estimates10_000 <- rep(0,10000)

estimates10_000 <- for (i in 1:10000) {
  estimates10_000[i] <- craps()
}

```

*3. Use your vector (of 100, then 1000, then 10 000) to provide a 95% confidence interval for the estimate of winning, e.g.: `t.test(estimates)`*

